\bf Nonparametric Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test for Marginal Distributions of Stationary Time Series

### 2024-03-24 manuscript/manuscript.tex
more references

The Kolmogorov-Smirnov (KS) test is a useful goodness-of-fit statistic. 
Let $X_1, ..., X_n$ be a random sample of size~$n$ from some continuous
distribution and the null hypothesis $H_0$ be that $X_i$'s follow the
hypothesized distribution~$F$.
If we let $F_n(t) = \sum_{i=1}^n I(X_i \le t) / n$ be the empirical cumulative
distribution function of the sample, where $I(\cdot)$ is the indicator
function, the KS test statistic takes the form
\begin{equation}
  \label{eq:ks_standard}
  T_n = \sqrt{n} \sup_x | F_{n}(x) - F(x) |.
\end{equation}
As sample size $n\to \infty$, the distribution of $T_n$ converges to that of the
absolute value of standard Brownian bridge, which is known as the Kolmogorov
distribution \citep{stephens1974edf}. The distribution function can be
accurately evaluated in modern statistical software
\citep{marsaglia2003evaluating}. The KS test can be applied to a variety of
fields. Examples are cosmic microwave background 
radiation \citep{naess2012application},
count rate of redioactive data??? \citep{aslam2020introducing},
gear condition monitoring \citep{andrade2001gear}, and
images of breast cancer tumors \citep{demidenko2004kolmogorov}.

Despite its widespread use, the KS test can be 
misapplied when its foundational assumptions are overlooked. The KS test 
assumes the data are independently and identically distributed (i.i.d.) and 
that the hypothesized distribution is continuous and fully specified without 
the need for parameter estimation. \citet{zeimbekakis2022misuses} examines 
common misapplications of the one-sample KS test, notably the inappropriate
use when the hypothesized distribution contains unspecified parameters.
For this situation, \citet{zeimbekakis2022misuses} proposed a parametric
bootstrap, where bootstrap samples of the test statistics are constructed from
samples generated from the fitted hypothesized distribution.
Alternatively, \citet{babu2004goodness} addresses this issue through the
application of basic non-parametric bootstrap methods.
Another commonly encountered situation is when the data are serially
dependent.
\citet{zeimbekakis2022misuses} proposed a semi-parametric bootstrap
where the serial dependence structure is modeled via a working serial copula.
A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semi-parametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for.
\jy{This reference is an example of abuse in time series settings:
Tuncer, O., Ates, E., Zhang, Y., Turk, A., Brandt, J., Leung, V. J., Egele, M., and Coskun, A. K. (2019), “Online Diagnosis of Performance Variation in HPC Systems Using Machine Learning,” IEEE Transactions on Parallel and Distributed Systems, 30, 883–896.
}

@@

First, let us consider how the KS statistic is typically computed for a sample
with no dependence structure.
Let $\hat\theta_n$ denote the fitted parameters for the hypothesized 
distribution fitted onto $X_t$, and let 
$F_n$ denote the empirical distribution function based on $X_1,...,X_n$
Let
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)).
\end{equation*}
Then, the
goodness of fit statistic is 
\begin{equation*}
T_n := \sup_x|Y_n(x; \hat\theta)|.
\end{equation*}

@@

Let us first consider the case where $X_i$'s are independent. Denote by
$F^{(b)}_n$ the empirical distribution of the $b$th bootstrap sample and let
$\hat\theta^{(b)}_n$ be the parameter estimate based on the $b$th bootstrap 
sample. 
Using the bootstrap (asymptotic) theory, we can approximate the distribution of
$\sqrt{n}(F_n(x) - F(x))$ and $\sqrt{n}(F(x; \hat\theta_n) - F(x))$
by that of $\sqrt{n}(F^{(b)}_n(x) - F_n(x))$ and
$\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; \hat\theta_n))$, respectively.
Therefore, if we define
\begin{align*}
Y^{(b)}_n(x) &= \sqrt{n}(F^{(b)}_n(x) - F_n(x)) - 
               \sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; \hat\theta_n)) \\
             &= \sqrt{n}(F^{(b)}_n(x) - F(x; \hat\theta^{(b)}_n)) - 
               \sqrt{n}(F_n(x) - F(x; \hat\theta_n)),
\end{align*}
then $T^{(b)}_n := \sup_x|Y^{(b)}_n(x)|$ is the bootstrap statistic that is 
expected to approximate the distribution of $T_n$. We note that the term
$\sqrt{n}(F_n(x) - F(x; \hat\theta_n))$ is exactly the bias term considered in 
\citet{babu2004goodness}.

@@

In this simulation study, we aim to show that the test correctly fails to 
reject $H_0$ under the null
distribution, and correctly rejects $H_0$ for an alternate distribution. If
both of these things are true, we can state that the method works
reasonably well.

@@

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$ with Kendall's
\jy{Need to state how serial dependence were introduced, like you did in the
  first paper: transforming a normal series.}
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
marginal distributions.
To generate the samples to which our
method would be applied, we simulated a time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(W_t)],
\end{equation*}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the $\Gamma(8, 1)$
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are $\{-0.607, 
diverges, -0.287, 0, 0.339, 0.668, 0.911\}$.
\mc{my explanation needs work}

@@

The Kolmogorov-Smirnov test provides a way for researchers in a broad array
of applications to compare the distributions of populations. The
method proposed in this paper offers a variant of the test that can
be utilized for serially dependent data without specifying parameters.
Using simulation, we have shown that given a large enough sample of a time 
series, the block 
bootstrap KS test can appropriately fail to reject the null hypothesis when the
series follows the null distribution. In addition, when the marginal
distribution is not
the one hypothesized, the test is powerful. It requires a larger sample size 
than basic bootstrap to avoid either type I or type II errors. Unsurprisingly,
the method performs better as the temporal dependence gets weaker. Through
simulation studies with Normal and Gamma-distributed data, we have shown that
that the method is robust. We have also 
demonstrated possible applications of this method to financial data.
Through comparison with \citep{babu2004goodness}'s non-parametric
bootstrap KS test and \citet{zeimbekakis2022misuses}'s parametric bootstrap
KS test, we have shown that our method can reach different conclusions which
would perhaps be missed if serial dependence was neglected and another
method was chosen.
This method can be used for any study where the goal is
to see if a serially dependenttime series follows a hypothesized distribution and the parameters
are unknown or unspecified. Future studies could investigate
the method's performance on different marginal distributions and
different dependence structures. One could also apply the method showcased
in this paper to fields like earthquake prediction and astronomy.

>>>>

The Kolmogorov-Smirnov (KS) test is a useful goodness-of-fit statistic. 
Let $X_1, ..., X_n$ be a random sample of size~$n$ from some continuous
distribution and the null hypothesis $H_0$ be that $X_i$'s follow the
hypothesized distribution~$F$.
If we let $F_n(t) = \sum_{i=1}^n I(X_i \le t) / n$ be the empirical cumulative
distribution function of the sample, where $I(\cdot)$ is the indicator
function, the KS test statistic takes the form
\begin{equation}
  \label{eq:ks_standard}
  T_n = \sqrt{n} \sup_x | F_{n}(x) - F(x) |.
\end{equation}
As sample size $n\to \infty$, the distribution of $T_n$ converges to that of the
absolute value of standard Brownian bridge, which is known as the Kolmogorov
distribution \citep{stephens1974edf}. The distribution function can be
accurately evaluated in modern statistical software
\citep{marsaglia2003evaluating}. The KS test can be applied to a variety of
fields. Examples are cosmic microwave background 
radiation \citep{naess2012application},
count rate of radioactive data \citep{aslam2020introducing},
gear condition monitoring \citep{andrade2001gear},
images of breast cancer tumors \citep{demidenko2004kolmogorov},
and financial markets \citep{lux2001turbulence}.

Despite its widespread use, the KS test can be 
misapplied when its foundational assumptions are overlooked. The KS test 
assumes the data are independently and identically distributed (i.i.d.) and 
that the hypothesized distribution is continuous and fully specified without 
the need for parameter estimation. \citet{zeimbekakis2022misuses} examines 
common misapplications of the one-sample KS test, notably the inappropriate
use when the hypothesized distribution contains unspecified parameters.
For this situation, \citet{zeimbekakis2022misuses} proposed a parametric
bootstrap, where bootstrap samples of the test statistics are constructed from
samples generated from the fitted hypothesized distribution.
Alternatively, \citet{babu2004goodness} addresses this issue through the
application of basic non-parametric bootstrap methods.
Another commonly encountered situation is when the data are serially
dependent. The KS test is frequently misused without accounting for the
assumption of independence. In an investigation of performance variation in
high-performance computing systems, \citet{tuncer2019ieee} did not describe
how or if they accounted for serial dependence when applying the two-sample
KS test.
\citet{zeimbekakis2022misuses} proposed a semi-parametric bootstrap
where the serial dependence structure is modeled via a working serial copula.
A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semi-parametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for.

\jy{This reference is an example of abuse in time series settings:
Tuncer, O., Ates, E., Zhang, Y., Turk, A., Brandt, J., Leung, V. J., Egele, M., and Coskun, A. K. (2019), “Online Diagnosis of Performance Variation in HPC Systems Using Machine Learning,” IEEE Transactions on Parallel and Distributed Systems, 30, 883–896.
}
\mc{addressed}

@@

\subsection{Review of Kolmogorov-Smirnov Test}

First, let us review how the KS statistic is typically computed for a sample
with no serial dependence and specified parameters.
Let $\hat\theta_n$ denote the fitted parameters for the hypothesized 
distribution fitted onto $X_t$, and let 
$F_n$ denote the empirical distribution function based on $X_1,...,X_n$.
Let
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)).
\end{equation*}
Then, the
goodness of fit statistic is 
\begin{equation*}
T_n := \sup_x|Y_n(x; \hat\theta)|.
\end{equation*}

@@

\subsection{Non-Parametric Basic Bootstrap}

Let us then consider the case where $X_i$'s are independent, but parameters
are unspecified. Denote by
$F^{(b)}_n$ the empirical distribution of the $b$th bootstrap sample and let
$\hat\theta^{(b)}_n$ be the parameter estimate based on the $b$th bootstrap 
sample. 
Using the bootstrap (asymptotic) theory, we can approximate the distribution of
$\sqrt{n}(F_n(x) - F(x))$ and $\sqrt{n}(F(x; \hat\theta_n) - F(x))$
by that of $\sqrt{n}(F^{(b)}_n(x) - F_n(x))$ and
$\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; \hat\theta_n))$, respectively.
Therefore, if we define
\begin{align*}
Y^{(b)}_n(x) &= \sqrt{n}(F^{(b)}_n(x) - F_n(x)) - 
               \sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; \hat\theta_n)) \\
             &= \sqrt{n}(F^{(b)}_n(x) - F(x; \hat\theta^{(b)}_n)) - 
               \sqrt{n}(F_n(x) - F(x; \hat\theta_n)),
\end{align*}
then $T^{(b)}_n := \sup_x|Y^{(b)}_n(x)|$ is the bootstrap statistic that is 
expected to approximate the distribution of $T_n$. We note that the term
$\sqrt{n}(F_n(x) - F(x; \hat\theta_n))$ is exactly the bias term considered in 
\citet{babu2004goodness}.

\subsection{Non-Parametric Block Bootstrap}

@@

In this simulation study, we aim to show that under the null
distribution, our test rejects the $H_0$ at the significance level, 
and that under the alternate distribution, our test rejects the $H_0$ at a high
rate. If
both of these things are true, we can state that the method works
reasonably well.

@@

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$ with Kendall's
\jy{Need to state how serial dependence were introduced, like you did in the
  first paper: transforming a normal series.}
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
marginal distributions.
To generate the samples to which our
method would be applied, we simulated a time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(W_t)],
\end{equation*}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the $\Gamma(8, 1)$
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are $\{-0.8837545, 
-0.686256, -0.3887456, -0.03336129, 0.3265704, 0.6389938, 0.8610347\}$.
\mc{my explanation needs work}

@@

The Kolmogorov-Smirnov test provides a way for researchers in a broad array
of applications to compare the distributions of populations. The
method proposed in this paper offers a variant of the test that can
be utilized for serially dependent data without specifying parameters.
Using simulation, we have shown that given a large enough sample of a time 
series, the block 
bootstrap KS test can appropriately fail to reject the null hypothesis when the
series follows the null distribution. In addition, when the marginal
distribution is not
the one hypothesized, the test is powerful. It requires a larger sample size 
than basic bootstrap to avoid either type I or type II errors. Unsurprisingly,
the method performs better as the temporal dependence gets weaker. Through
simulation studies with Normal and Gamma-distributed data, we have shown that
that the method is robust. We have also 
demonstrated possible applications of this method to financial data.
Through comparison with \citep{babu2004goodness}'s non-parametric
bootstrap KS test and \citet{zeimbekakis2022misuses}'s parametric bootstrap
KS test, we have shown that our method can reach different conclusions which
would perhaps be missed if serial dependence was neglected and another
method was chosen.
This method can be used for any study where the goal is
to see if a serially dependent time series follows a hypothesized distribution
and the parameters
are unknown or unspecified. Future studies could investigate
the method's performance on different marginal distributions and
different dependence structures. Additionally, one could propose a 
similar method
for comparing two serially dependent methods. One could also apply the method
showcased
in this paper to fields like earthquake prediction and astronomy.
